---
title: "Eviction chorepleth map"
author: "Zhiyu Wei"
date: 2024-10-14
output: github_document
---

#### Setup

```{r setup, include = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(knitr)
library(sf)
library(lubridate)
library(stringr)
library(leaflet)
library(shiny)
library(leaflet.extras)
library(htmlwidgets)
library(rgeoda)
library(RColorBrewer)
library(classInt)
```


```{r load evictions data}
evic = read.csv("./data/evictions.csv")

uhf = st_read("./UHF_42_DOHMH_2009/UHF_42_DOHMH_2009.shp")
```

```{r rename and clean}
evic =janitor::clean_names(evic)

uhf = janitor::clean_names(uhf)

# change date format
evic$executed_date <- mdy(evic$executed_date)

# add year to data 
evic = evic |>
  mutate(year = format(executed_date, "%Y"))
```

#### EDA

```{r Eviction EDA}
# eviction by boroughs
evic |>
  group_by(borough) |>
  summarize(Num_evic = n())|>
  knitr::kable(col.names = c("Borough", "Number of Evictions"))

# evictions by year
evic |>
  group_by(year) |>
  summarize(num_evic = n()) |>
  knitr::kable(col.names = c("Year", "Number of Evictions"))

# by both
evic |>
  group_by(year, borough) |>
  summarize(num_evic = n()) |>
  knitr::kable(col.names = c("Year", "NYC boroughs", "Number of Evictions"))
```


```{r bar graphs}
# by borough
evic |>
  ggplot(aes(x = borough))+
  geom_bar() +
  labs(x ="NYC Boroughs",
        y = "Number of Evictions")

# by year
evic |>
  filter(!(year == 2025)) |>
  ggplot(aes(x = year))+
  geom_bar() +
  labs(x = "Year",
        y = "Number of Evictions")

# Stacked up by borough
evic|>
  group_by(year) %>%
  count(borough) %>%
  ggplot(aes(x = year, y = n, fill = borough)) +
  geom_col() +
  labs(x = "Year",
       y = "Number of Evictions")
  
```

#### write out the csv
```{r csv for evictions}
write.csv(evic, "./data/evictions_for_Qgis.csv")

# 2022 
evic_22 = evic |>
  filter(year == 2022)

write.csv(evic_22, "./data/eviction_2022.csv")
```

#### counting points
```{r point count}
#save coordinates as geometry
evic = evic |>
  na.omit()|>
  st_as_sf(coords = c("longitude","latitude"), crs = 4326)

# transform coordinate reference system
if (!st_crs(evic) == st_crs(uhf)) {

# Transform evic to match uhf

  uhf_transform <- st_transform(uhf, st_crs(evic)) 
} 

# loop to create single dataset for counts in every single year:
years <- 2017:2024

for (year in years) {
  assign(
    paste0("uhf_evic_", year),
    uhf_transform |>
      na.omit() |>
      mutate(
        counts = lengths(st_intersects(geometry, evic |> filter(year == !!year))),
        year = year
      )
  )
}

# bind all years together
uhf_evic = rbind (uhf_evic_2019, uhf_evic_2020, uhf_evic_2021, uhf_evic_2022, uhf_evic_2023, uhf_evic_2024) |>
  group_by(borough, uhf_neigh) |>
  summarize(counts = mean(counts, na.rm=TRUE))

st_write(uhf_evic, "./data/uhf_evic.shp")

# check coordinate referencce system
st_crs(uhf_evic)

```

#### Using UHF

```{r uhf}
# Mapping with UHF polygons (see the 42 neighborhoods)
ggplot()+
  geom_sf(data = uhf)
```

#### Eviction Leaflet

Layer 1: Display dataset with pop-up

```{r Display dataset with pop-up}

# Base tiles
leaflet() |> 
  addProviderTiles("CartoDB", group = "CartoDB")

#add multiple base maps
leaflet() |> 
  addProviderTiles("CartoDB", group = "CartoDB") |> 
  addProviderTiles("CartoDB.DarkMatterNoLabels", group = "CartoDB - Black") 


# Control layer
leaflet() |>  
  addProviderTiles("CartoDB", group = "CartoDB") |> 
  addProviderTiles("CartoDB.DarkMatterNoLabels", group = "CartoDB - Black") |> 
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black")) 
######

# Specify location of interest
leaflet() |>  
  addProviderTiles("CartoDB", group = "CartoDB") |> 
  addProviderTiles("CartoDB.DarkMatterNoLabels", group = "CartoDB - Black") |> 
  setView(lng = -74.006, lat = 40.7128, zoom = 10) |>  # Coordinates of NYC
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black")) 


# Assign it to an object
leaflet() |>  
  addProviderTiles("CartoDB", group = "CartoDB") |> 
  addProviderTiles("CartoDB.DarkMatterNoLabels", group = "CartoDB - Black") |> 
  setView(lng = -74.006, lat = 40.7128, zoom = 10) |>  # Coordinates of NYC
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black")) -> map_tiles
```


#### Add choropleth layer 
### need to tailor for each year and then make different layer for each

##### no need to run this chunk, if different method of statistics is needed, then run this chunk to modify codes. 
``` {R choropleth}
# palatte for total evictions
pal_uhf_evic <- colorNumeric("Reds", 
                     domain = uhf_evic$counts,
                     reverse = FALSE)



datasets <- lapply(years, function(year) get(paste0("uhf_evic_", year)))

# Generate color palettes dynamically
palettes <- lapply(datasets, function(data) {
  colorNumeric("Reds", domain = data$counts, reverse = FALSE)
})

# Assign palettes to variables dynamically
names(palettes) <- paste0("pal_uhf_evic_", years)
list2env(palettes, envir = .GlobalEnv) 
  
for (year in years) {
  # Get the dataset and palette for the current year
  current_data <- get(paste0("uhf_evic_", year))
  current_palette <- get(paste0("pal_uhf_evic_", year))
  
  # Add polygons and legend for the current year
  map_tiles <- map_tiles |>
    addPolygons(
      data = current_data,  # Dataset for the current year
      group = paste0("Evictions ", year),
      color = "white",
      weight = 1,
      fillColor = ~current_palette(counts),  # Corrected palette usage
      fillOpacity = 1,
      label = ~uhf_neigh
    ) }

# Add the total evictions layer and legend
map_tiles <- map_tiles |>
  addPolygons(
    data = uhf_evic,
    group = "Total Evictions",
    color = "white",
    weight = 1,
    fillColor = ~pal_uhf_evic(counts),
    fillOpacity = 1,
    label = ~uhf_neigh
  ) 

# Add layer controls and hide all groups initially
map_tiles = map_tiles |>
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black"),
                   overlayGroups =  c("Total Evictions", paste0("Evictions ", years)),  
                   # Add all layers to control
    options = layersControlOptions(collapsed = FALSE)) |>
    hideGroup(c("Total Evictions", paste0("Evictions ", years)))  
# Hide all layers and legends initially

```

```{r generate quantiles}
# Generate equal quantiles (5 classes)

# Loop through years to create quantile breaks for each year
for (year in years) {
  assign(
    paste0("quantile_breaks_", year),
    classIntervals(
      get(paste0("uhf_evic_", year))$counts,  # Dynamically access the dataset
      n = 5,  # Number of quantile breaks
      style = "quantile"  # Quantile style
    )
  )
}

quantile_breaks <- classIntervals(uhf_evic$counts, n = 5, style = "quantile")



pal_uhf_evic2 <- colorBin(palette = "Reds",             
                      domain = uhf_evic$counts,   
                      bins = c(18.50,  79.43333, 125.90, 177.93333, 316.23333, 630.16667),        
                      reverse = FALSE)

# Loop through years and generate color bins
for (year in years) {
  # Retrieve the quantile breaks for the current year
  quantile_breaks <- get(paste0("quantile_breaks_", year))$brks
  
  # Create the color bin palette dynamically
  assign(
    paste0("pal_uhf_evic_2_", year),
    colorBin(
      palette = "Reds",
      domain = get(paste0("uhf_evic_", year))$counts,  # Counts for the year
      bins = quantile_breaks,  # Quantile breaks
      reverse = FALSE
    )
  )
}

# Add labels and line breaks to total evictions
map_tiles = map_tiles |> 
  addPolygons(data = uhf_evic,
              group = "Total Evictions",
              color = "white",
              weight = 1,
              fillColor = ~pal_uhf_evic2(counts),
              fillOpacity = 1,
              label = ~uhf_neigh,
              popup = ~paste(
      "<b>Borough:</b>", borough, "<br>",
      "<b>Neighborhood:</b>", uhf_neigh, "<br>",
      "<b>Number of Evictions:</b>", counts, "<br>")) |> 
  addLegend(pal = pal_uhf_evic2,
            values = uhf_evic$counts,
            group = "Total Evictions",
            title = "Total Evictions",
            position = "bottomright") 

########

# Loop through years to dynamically add polygons and legends

for (year in years) {
  # Get the dataset and palette for the current year
  current_data <- get(paste0("uhf_evic_", year))
  current_palette <- get(paste0("pal_uhf_evic_2_", year))
  
  # Add polygons, legend, and pop-ups for the current year
  map_tiles <- map_tiles |>
    addPolygons(
      data = current_data,
      group = paste0("Evictions ", year),  # Group for this year's layer
      color = "white",  # White border for polygons
      weight = 1,
      fillColor = ~current_palette(counts),  # Use the current year's palette
      fillOpacity = 1,
      label = ~uhf_neigh,
      popup = ~paste(
        "<b>Borough:</b>", borough, "<br>",
        "<b>Neighborhood:</b>", uhf_neigh, "<br>",
        "<b>Number of Evictions:</b>", counts, "<br>",
        "<b>Year:</b>", year
      )
    ) |>
    addLegend(
      pal = current_palette,  # Use the current year's palette for the legend
      values = current_data$counts,
      title = HTML(paste0("<span style='font-size:14px;'>Number of Evictions ", year, "</span>")),
      position = "bottomleft",
      group = paste0("Evictions ", year)  # Link the legend to the corresponding layer
    )
}

# Add layer controls and hide all groups initially
map_tiles = map_tiles |>
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black"),
                   overlayGroups =  c("Total Evictions", paste0("Evictions ", years)),  
                   # Add all layers to control
    options = layersControlOptions(collapsed = FALSE)) |>
    hideGroup(c("Total Evictions", paste0("Evictions ", years)))  
# Hide all layers and legends initially



######## test for debugging

uhf_evic_2017_test <- uhf_evic_2017 |> 
  mutate(popup_content = paste(
    "<b>Borough:</b>", borough, "<br>",
    "<b>Neighborhood:</b>", uhf_neigh, "<br>",
    "<b>Number of Evictions:</b>", counts, "<br>",
    "<b>Year:</b>", 2017
  ))



map_tiles_2017_test <- leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = uhf_evic_2017_test,
    color = "white",
    weight = 1,
    fillColor = ~pal_uhf_evic_2_2017(counts),
    fillOpacity = 1,
    label = ~uhf_neigh,
    popup = ~popup_content
  )

```

## Layer 2: Display statistically significant High-High and Low-Low clusters

```{r clusters layers}
#Load bivariate shape file
clusters <-  st_read("./data/US_CTY_PreventativeHealthExp.shp") |>
  filter(LISA_CL == 1 | LISA_CL ==2) |>
  mutate(cluster = if_else(LISA_CL == 1, "High - High", "Low - Low", NA_character_))

#re project
clusters <- st_transform(clusters, crs = 4326)

# Add the clusters 
pal_cluster <- colorFactor(palette = c("darkred","darkRedse"),
                           domain = clusters$cluster)
map_tiles |> 
  addPolygons(data = uhf_evic,
              color = "black",
              weight = 1,
              fillColor = ~pal_uhf_evic2(counts),
              fillOpacity = 1,
              label = ~uhf_neigh,
              popup = ~paste(
                "<div style='font-family: Roboto; font-size: 12px; line-height: 1.4;'>",
                "<div style='font-size: 12px; font-weight: bold; color: #d9534f;'><b>Neighborhood:</b> ",  uhf_neigh,"<br>",
                      "<b>Expenditures on Preventative Health Programs

:</b>", counts,"<br>",
                      "<b>Population who reside in rural places:</b>", PCT_RUR,"%<br>",
                      "<b>Population falls under the poverty line:</b>", PCT_POV,"%<br>",
                      "<b>Population with BA degrees:</b>", PCT_BA,"%<br>",
                      "<b>Population who are White:</b>", PCT_WH,"%<br>" ,
                "</div>")) |> 
  addLegend(pal = pal_uhf_evic2,
            values = uhf_evic$counts,
            title = "Number of Evictions",
            position = "bottomright") |> 
  addPolygons(data = clusters,
             color = ~pal_cluster(cluster),
             weight = 2,
             fillColor = NA,
             popup = ~paste(cluster, "<br>"),
             group = "Clusters") |> 
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black"),
                   overlayGroups = c("Number of Evictions", "Clusters"), 
                   options = layersControlOptions(collapsed = FALSE))

#### Prepare map for export ----
map_tiles |> 
  addPolygons(data = uhf_evic,
              color = "black",
              weight = 1,
              fillColor = ~pal_uhf_evic2(counts),
              fillOpacity = 1,
              label = ~uhf_neigh,
              popup = ~paste(
                "<div style='font-family: Roboto; font-size: 12px; line-height: 1.4;'>",
                "<div style='font-size: 12px; font-weight: bold; color: #d9534f;'><b>Neighborhood:</b> ",  uhf_neigh,"<br>",
                      "<b>Expenditures on Preventative Health Programs

:</b>", counts,"<br>",
                      "<b>Population who reside in rural places:</b>", PCT_RUR,"%<br>",
                      "<b>Population falls under the poverty line:</b>", PCT_POV,"%<br>",
                      "<b>Population with BA degrees:</b>", PCT_BA,"%<br>",
                      "<b>Population who are White:</b>", PCT_WH,"%<br>" ,
                "</div>")) |> 
  addLegend(pal = pal_uhf_evic2,
            values = uhf_evic$counts,
            title = "Number of Evictions",
            position = "bottomright") |> 
  addPolygons(data = clusters,
             color = ~pal_cluster(cluster),
             weight = 2,
             fillColor = NA,
             popup = ~paste(cluster, "<br>"),
             group = "Clusters") |> 
  addLayersControl(baseGroups = c("CartoDB", "CartoDB - Black"),
                   overlayGroups = c("Number of Evictions", "Clusters"), 
                   options = layersControlOptions(collapsed = FALSE)) -> Section4_map

# Save interactive maps 
saveWidget(eviction,
           file = "eviction_choropleth.html")


```


